---
title: "Identifying Best Date to Find Chicken of the Woods"
author: "Alex Duman"
output: html_document
date: "`r format(Sys.time(), '%B %d, %Y')`"
---
**DISCLAIMER:** Consuming wild mushrooms, plants or other material is inherently risky for health reasons as well as potentially damaging to local ecosystems. Please do not consume anything you are not 100% confident identifying and be respectful, follow local regulations and  sustainable harvesting practices if you do chose to forage (e.g., don't take something if it is the only one, don't take more than you need, get permission to harvest on private property, etc.).


**Background:** An estimated 3.5 to 5.8 billion people collect non-timber forest products globally, this includes harvesting wild herbs, fruits, fungi, and more [[1](https://doi.org/10.1016/j.forpol.2021.102659)]. 

Approximately 1-in-4 people surveyed within the New England region of the United States of America, reported harvesting some non-timber forest products within the past 5 years [[2](https://doi.org/10.1111/j.1475-4762.2008.00794.x)].

This analysis focuses on *Laetiporus*, a genus of edible mushrooms that decompose dead wood and are found around the world. This genus of fungi is better known by its common name, 'chicken of the woods' due to its distinct flavor and texture which resembles chicken.

**Problem:** Newer foragers, or seasoned foragers who have not encountered a chicken of the woods mushroom may not know when the best time to look for these iconic mushrooms.

**Objective:** We'll leverage generalized additive models (GAMs) to help us assess the best day of the year to find a chicken of the woods mushroom in a particular region using available historic data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First I'll load all the necessary packages and the data I queried directly from [iNaturalist](https://www.inaturalist.org/). These data include research grade observations (which require at least 3 separate species identification from community members) of all four species within the genus *Laetiporus* as well as one subspecies.
```{r setup-2, warning=FALSE}
# Set working directory
setwd('C:/Users/alexj/Downloads/ChickenoftheWoods')

# Load Packages
library("grid")
library("lubridate")
library("magick")
library("maps")
library("mapview")
library("mgcv")
library("sf")
library("tidyverse")
library("tigris")
options(tigris_class = "sf")
library("yardstick")

# Read in data sets
df1 <- read.csv("L_sulphureus_observations.csv")
df2 <- read.csv("L_cincinnatus_observations.csv")
df3 <- read.csv("L_conifericola_observations.csv")
df4 <- read.csv("L_gilbertsonii_observations.csv")

# Combine data into a single data set
df <- rbind(df1, df2, df3, df4) %>%
  select(id, latitude, longitude, observed_on, scientific_name, common_name)
rm(df1, df2, df3, df4) # remove individual data sets

# Check the structure of the primary data set
str(df)
```

We have `r nrow(df)`` rows or observations of chicken of the woods mushrooms.

The columns of our data set include:
**id** - the unique identifying number for each record.
**latitude** - geographic coordinate indicating North-South location relative to the
           equator (0 degrees) versus the North and South Poles (90 degrees).
**longitude** - geographic coordinate measuring East-West location from the prime 
            meridian (range: -180 to 180).
**observed_on** - the date the observation was made on iNaturalist.
**scientific_name** - the Latin taxonomic name of the species.
**common_name** - the plain English species name.

While latitude and longitude are helpful, we'll get more helpful location information using the `tigris` package, which contains the necessary information to convert latitude and longitude into US Census Bureau defined regions.

### Preparing Data
```{r preparing-data}
# Get date data
df <- df %>%
  mutate(observed_on = as.Date(observed_on),
         Day = day(observed_on), # Day of the month observation was made
         Month = month(observed_on), # Month of the year observation was made
         Year = year(observed_on), # Year of observation
         Weekend = ifelse(wday(observed_on) %in% c(1,7), TRUE, FALSE), # Whether observed on a weekend
         days_into_year = as.numeric(observed_on - as.Date(paste0(as.character(Year), "-01-01"))))

# Get states
coordinates <- st_as_sf(df, coords = c("longitude", "latitude"), crs = 4326, agr = "constant")
USstates <- states()
coordinates <- st_transform(coordinates, st_crs(USstates))
coords_USstate <- st_contains(USstates, coordinates) %>%
  as.data.frame()
coordinates_with_state <- st_join(coordinates, USstates) %>%
  select(id, REGION, STUSPS, NAME) %>%
  rename(ST = STUSPS, State = NAME, Region = REGION)

# Add state data to data set
df <- left_join(df, coordinates_with_state, by = join_by(id == id))  %>%
  mutate(Region = ifelse(State %in% c("Alaska", "Hawaii"), 5, Region))
rm(coordinates, USstates, coords_USstate, coordinates_with_state)

# Check updated data set
summary(df)
```
The newly added columns of our data set include:
**Day** - the numerical day of the month the observation was made,
**Month** - the numerical value of the month the observation was made,
**Year** - the year in which the observation was made,
**Weekend** - logical indicator for whether observation was made on a weekend,
**days_into_year** - the numerical number of days into the year the observation was 
                 made (range: 0 to 366, with leap year),
**Region** - indicates a geographic region of the United States (1 = Northeast, 
         2 = Midwest, 3 = South, 4 = Western Continental, and 5) Alaska & Hawaii),
**ST** - is the two-letter abbreviation for each state, and
**State** - the full name of the state.

Of particular note is that a large percent of observations (39.2%) were made on a weekend, which is higher than we would expect if people were just as likely to find a chicken of the woods on any given day of the week (e.g., 2/7 = 28.6%). One potential explanation is that people are more likely to observe a chicken of the woods mushroom on the weekend because they're more likely to go hiking or foraging on weekends when they have more free time (just a hypothesis, and one you could consider exploring further on your own).

We'll work with data at the level of the state since there tend to be a good amount of observations per state and this provides better estimates than for the entire region. However, before we do that we need to clean the data that didn't get labeled with a state.
## Clean Data
```{r check-missing-state-data}
# Check data with missing state names
df %>%
  filter(is.na(State)) %>%
  select(id, latitude, longitude) %>%
  arrange(id) %>%
print()
```
Because there were only 19 missing State values I manually checked which states related to each of the above points so we can update the data set accordingly. Checking these latitude and longitude values in [Google maps](https://www.google.com/maps), I found most of the coordinates were slightly off shore so we're assigning them to the nearest state. There were also two records that showed up just over the boarder in Canada. 

We'll assume these points were randomly generated because the observer can obscure the exact location information which randomly displays the location in a 0.2 by 0.2 degree rectangular grid (for more information see [iNaturalist's geoprivacy policy](https://www.inaturalist.org/pages/geoprivacy)) 
```{r update-states}
# Update State values for those missing a State name
df_States <- data.frame(id = c(34559282, 62719394, 66394994, 94289848, 95431170,
                               95998343, 96326401, 100796907, 137207910, 138267408, 
                               139033843, 142109170, 181940907, 187144285, 187706373, 
                               204219413, 231346285, 233114679, 237047841),
                        State_updated = c("California", "California", "California", "California", "Rhode Island", 
                                  "New York", "California", "California", "California", "California",
                                  "California", "California", "California", "New York", "California",
                                  "New York", "Massachusetts", "California", "California"))

# Update state
df <- df %>%
  left_join(df_States, by = "id") %>%
  mutate(State = ifelse(is.na(State_updated), State, State_updated)) %>%
  select(-State_updated)

# Remove extra data frame now that we no longer need it
rm(df_States)

# Check data with missing state names
df %>%
  filter(is.na(State)) %>%
  select(id, latitude, longitude) %>%
  arrange(id) %>%
print()
```
We no longer have data with missing State data! Next, I'll make a few visualizations to better understand the data.

## Exploratory Data Analysis
```{r}
mapview(df, xcol = "longitude", ycol = "latitude", zcol = "scientific_name", crs = 4269, grid = FALSE)
```
Here we see that many observations are made within the Eastern United States, along the West coast, and in Southern Alaska as well as Hawaii.

```{r}
# Check when observations were made
df %>%
  group_by(observed_on) %>%
  summarize(Observations = n()) %>%
ggplot(aes(x=observed_on, y=Observations)) + 
  geom_line(color="black") + 
  labs(title = "Observations through time",
       x = "Date", y = "Number of Observations") + 
  theme_classic()
```
It looks like there are relatively consistent annual cycles, at least over the past decade. However, because iNaturalist didn't debut until March of 2008, we'll be conservative and only rely on data observed between January 1st, 2010 and December 31st, 2023. I've cut off the dataset at the end of 2023 because I'd like to use data from a complete year to have a sense of when the most mushrooms have historically been found.
```{r}
# Get original sample size
N_obs_orig <- nrow(df)

# Filter down to years of interest
df <- filter(df, Year >= 2010 & Year <= 2023)

# Check the filtered data
glimpse(df)
```
Now we have `r nrow(df)` observations which is still `r paste0(as.character(round(nrow(df)/N_obs_orig*100,1)),"%")`` of the original data we imported.

```{r visualize-temporal-distributions}
# Check temporal distribution of observations stratified by species
ggplot(df, aes(x = Month, fill = scientific_name)) + 
  geom_bar() + 
  scale_fill_manual(values = c("#5b0fa1", "#2a68a8", "#2ea6a7", "#88e458", "#ffef00")) + 
  facet_wrap(~scientific_name, nrow=5) + 
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12),
                     labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  theme_classic()
```
We can see each species has a different distribution of observations but
across the board most observations are made between July and October. It is also clear that some species are observed at much higher rates (e.g., 1,000's of times per month vs 100's of times per month), could be a result of species abundance or of population density in areas where certain species are naturally occurring. 

```{r}
# Visualize observations by day of the year
all_days <- data.frame(days_into_year = 0:365)

df_plot <- df %>%
  group_by(days_into_year) %>%
  summarize(observations = n()) %>%
  right_join(all_days, by = "days_into_year") %>%
  replace_na(list(observations = 0))

ggplot(df_plot, aes(x = days_into_year, y = observations)) + 
  geom_line() +
  geom_vline(xintercept = c(30, 59, 90, 120, 151, 181, 212, 243, 273, 304, 334), linetype = "dotted") + 
  annotate("text", x=c(15, 45, 75, 105, 135, 165, 195, 228, 260, 290, 320, 350), 
           y=1.1*max(df_plot$observations),
           label = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
           color="blue") + 
  labs(title = "Observations of Chicken of the Woods Over Time", 
       x = "Days into Year", y = "Observations",
       color = "Year") + 
  scale_x_continuous(breaks=c(0, 365)) +
  theme_classic()
```
When I combine data across all 13 years, it looks like most observations are made around late September and early October. However, while the above visualization while helpful for seeing the overall trend across the entire U.S. does not accurately account for regional variation driven by the climate, species found in that area, or population density which may result in certain states with larger populations driving the overall U.S. observation rate. 

I'll use a Generalized Additive Model (GAM) to help create a smoothed annual trend so we can better estimate the best day to find chicken of the woods mushrooms within each state. which handle non-linear relationships well and interactions between predictors, to help estimate the date with the most observations within for state (with observations). I've decided to use a GAM because this model type handle non-linear relationships well and interactions between predictors, which will help us estimate annual cycles in chicken of the woods mushroom observations within each state.

# Prepare Data for Modeling
```{r}
all_days_states <- all_days %>%
  cross_join(data.frame(State = sort(unique(df$State)))) %>%
  arrange(State, days_into_year)

# Now aggregate the data into total counts per day
df_model <- df %>%
  group_by(days_into_year, State) %>%
  summarize(observations = n()) %>%
  right_join(all_days_states, by = c("days_into_year", "State")) %>%
  mutate(observations = ifelse(is.na(observations) | is.nan(observations), 0, observations)) %>%
  arrange(State, days_into_year, State) %>%
  mutate(State = factor(State))

# GAM model incorporating days into year and the state
gam_model <-gam(observations ~ State + 
                  s(days_into_year, by = State,
                    bs = "cc"), # a cyclic cubic regression spline to ensure end points match
                 data = df_model)

# check summary of GAM model
summary(gam_model)
```
This model explains 90% of the variation in the number of chicken of the woods observations using the predictors, which is exceptionally high (R-square value over 0.90).

Now we'll use this model to make predictions about the number of observations on a given day of the year.
```{r}
# Get predicted values
df_model$predicted <- predict(gam_model, newdata = df_model, type = "response", se.fit = FALSE)
df_model$pred_se <- predict(gam_model, newdata = df_model, type = "response", se.fit = TRUE)$se.fit
```

### Make Predictions for Each State for which we had data
```{r}
# Setup best_day as a data frame
df_BestDay <- data.frame(matrix(NA, nrow = length(unique(df$State)), ncol = 3))
colnames(df_BestDay) <- c("days_into_year", "State", "Date")

# get predicted values from model for each state
for (i in 1:length(unique(df$State))){
  state <- sort(unique(df$State))[i]
  
  # Ensure we only make predictions for states we have data
  if(!is.na(state)){ 
    # Get date associated with most observations for each state
    df_BestDay$days_into_year[i] <- df_model$days_into_year[df_model$State == state & df_model$predicted==max(df_model$predicted[df_model$State==state])]
    df_BestDay$State[i] <- state
    best_date <- as.Date(paste0(format(Sys.Date(), "%Y"), "-01-01")) + df_BestDay$days_into_year[i]
    day <- as.character(as.numeric(format(best_date, "%d")))
    month <- format(best_date, "%B")
    suffix <- ifelse(day %in% c("11", "12", "13"), "th", 
                     ifelse(substring(day, nchar(day), nchar(day))=="1", "st",
                            ifelse(substring(day, nchar(day), nchar(day))=="2", "nd",
                                   ifelse(substring(day, nchar(day), nchar(day))=="3", "rd", "th"))))
    df_BestDay$Date[i] <- paste0(month, ' ', day, suffix)
    df_BestDay$Date_short[i] <- paste0(month(best_date), "/", day(best_date))
  }
}
```

### Visualizing Individuals States
I've just done a quick check for the three states I've lived in most recent: New York, California, and Illinois.
```{r}
# Define a State Visualization Function
viz_state <- function(df, df_BestDay, state){
  # This function helps visualize an individual state's chicken of the woods 
  # observations over the course of the year.
  
  # INPUTS
  #  df - data frame containing the observations, predictions, and time to plot
  #  df_BestDay - data frame containing the best date information for each state
  #  State - a character string for the state you want to visualize
  
  # Output
  #  visualization of mushroom observations over time
  
  # Ensure only data for the sate of interest is included
  df <- df %>%
    filter(State == state)
  
  # Visualize
  ggplot(df) + 
    geom_line(aes(x=days_into_year, y=observations), color="black") +
    geom_line(aes(x=days_into_year, y=predicted), size = 1, color="blue") + 
    geom_ribbon(aes(x=days_into_year, ymin = predicted - 1.96*pred_se,
                    ymax = predicted + 1.96*pred_se), alpha = 0.3, fill="blue") +
    geom_vline(xintercept=df_BestDay$days_into_year[df_BestDay$State==state], color="red") + 
    scale_x_continuous(breaks=c(0, df_BestDay$days_into_year[df_BestDay$State==state], 365)) +
    annotate("text", x=df_BestDay$days_into_year[df_BestDay$State==state]-7, y=max(df$predicted[])/4, 
             label = df_BestDay$Date[df_BestDay$State==state], color="red", angle = 90) +
    labs(title=paste0("Best Date in ", state, " State to find Chicken of the Woods"),
         x = "Number of days into year",
         y = "Number of Mushroom Observations") + 
    theme_classic()
}

# Visualize Results for New York
viz_state(df_model, df_BestDay, "New York")
```
```{r}
# Visualize Results for California
viz_state(df_model, df_BestDay, "California")
```
```{r}
# Visualize Results for Illinois
viz_state(df_model, df_BestDay, "Illinois")
```

# Visualizing all States within a Region
```{r}
# Add Region back onto the data frame to create regional plots
df_Regions <- df %>%
  filter(!is.na(Region)) %>%
  group_by(Region, State) %>%
  summarize(count = n()) %>%
  select(Region, State) %>%
  arrange(State)

# Join regions onto model data and Best Day data
df_model <- df_model %>%
  left_join(df_Regions, by = "State")
df_BestDay <- df_BestDay %>%
  left_join(df_Regions, by = "State")

# Define a Regional Vizualization Function
viz_region <- function(df, df_BestDay, region, region_name){
  # This function helps visualize an individual state's chicken of the woods 
  # observations over the course of the year.
  
  # INPUTS
  #  df - data frame containing the observations, predictions, and time to plot
  #  df_BestDay - data frame containing the best date information for each state
  #  region - an integer value between 1 and 5 for the region of interest
  #  region_name - a character string for the region name (e.g., Northeast U.S.)
  
  # Output
  #  visualization of mushroom observations over time for each state within that region
  
  # Ensure only data for the sate of interest is included
  df <- filter(df, Region == region)
  df_BestDay <- filter(df_BestDay, Region == region)
  
  # Visualize
  ggplot(data = df) + 
    geom_line(aes(x=days_into_year, y=observations), color="black") +
    geom_line(aes(x=days_into_year, y=predicted), size = 1, color="blue") + 
    geom_ribbon(aes(x=days_into_year, ymin = predicted - 1.96*pred_se, ymax = predicted + 1.96*pred_se), 
                alpha = 0.3, fill="blue") +
    geom_vline(data = df_BestDay, aes(xintercept=days_into_year, color="red")) +  #df_pred_long$BestDay
    geom_text(data = df_BestDay, aes(x=days_into_year-7, y=max(df$predicted)/2, label = Date_short), color="red", angle = 90) +
    labs(title=paste0("Best Day to Find Chicken of the Woods in the ", region_name),
         x = "Number of days into year",
         y = "Number of Mushroom Observations") + 
    facet_wrap(~State) + 
    theme_classic()
}

# Plotting data for the Northeast U.S. (Region 1)
viz_region(df_model, df_BestDay, region=1, region_name="Northeast U.S.")
```
```{r}
# Plotting data for the Midwest (Region 2)
viz_region(df_model, df_BestDay, region=2, region_name="Midwest")
```
```{r}
# Plotting data for the Midwest (Region 3)
viz_region(df_model, df_BestDay, region=3, region_name="Southern U.S.")
```
```{r}
# Plotting data for the Midwest (Region 4)
viz_region(df_model, df_BestDay, region=4, region_name="Western Continental U.S.")
```
```{r}
# Plotting data for Alaska & Hawaii (Region 5)
viz_region(df_model, df_BestDay, region=5, region_name="Alaska & Hawaii")
```

# Create animated gif of observations over time
```{r}
# Get State data
us_states <- map_data("state")

# Weekly GIF Creation
plot_list <- list()
j <- 1
week <- (0:51)*7
for (i in 1:length(week)){
  days_into_yr <- sort(unique(df$days_into_year))[i]
 
    if (i==length(week)){
      df_plot <- df[df$Region !=5 & df$days_into_year >= week[i] & df$days_into_year <=365, ] # the last week may be a little long
    } else {
      df_plot <- df[df$Region !=5 & df$days_into_year >= week[i] & df$days_into_year < week[i] + 7, ]
    }
  
  # Getting date range for data  
  start_date <- as.Date("2024-01-01") + week[i]
  if (i==length(week)){
    end_date <- as.Date("2024-12-31")
  } else {
    end_date <- as.Date("2024-01-01") + week[i] + 6
  }
  
  # Determine whether whether two months are needed
  start_date2 <- format(start_date, "%b %d")
  if (month(start_date)==month(end_date)){
    end_date2 <- format(end_date, "%d")
  } else {
    end_date2 <- format(end_date, "%b %d")
  }
  len_s <- nchar(start_date2)
  len_e <- nchar(end_date2)
  start_suffix <- ifelse(start_date2[len_s-1:len_s] %in% c("11", "12", "13"), "th", 
                         ifelse(substring(start_date2, len_s, len_s)=="1", "st",
                                ifelse(substring(start_date2, len_s, len_s)=="2", "nd",
                                       ifelse(substring(start_date2, len_s, len_s)=="3", "rd", "th"))))
  end_suffix <- ifelse(end_date2[len_e-1:len_e] %in% c("11", "12", "13"), "th", 
                       ifelse(substring(end_date2, len_e, len_e)=="1", "st",
                              ifelse(substring(end_date2, len_e, len_e)=="2", "nd",
                                     ifelse(substring(end_date2, len_e, len_e)=="3", "rd", "th"))))
  formatted_date <- paste0(start_date2, start_suffix, " - ", end_date2, end_suffix)
  
  # Create a plot
  p <- ggplot()
  if (nrow(df_plot) > 0){
    # If observations were observed during time period
    p <-  p + 
      geom_point(data = df_plot, aes(x=longitude, y=latitude), color="orange", alpha=0.3, size=1)
  }
  p <- p + 
    geom_polygon(data = us_states, aes(x = long, y = lat, group = group), fill=NA, color="black") +
    theme_classic() + 
    labs(title = paste0("Chicken of the woods observations: ",formatted_date)) + 
    theme(axis.line = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.title = element_blank(),
          plot.title = element_text(color="blue", face="bold")) + 
    coord_fixed(1.3) # set aspect ratio 
  
  ggsave(filename = paste0("plot", as.character(i), ".png"), plot=p)
  plot_list[[j]] <- image_read(paste0("plot", as.character(i), ".png"))
  j <- j + 1
}

# Combine into GIF
gif <- image_animate(image_join(plot_list), fps = 2)
image_write(gif, "ChickenOfTheWoods_weekly.gif")

# Delete all the png files
png_files <- list.files(pattern = "*.png")
file.remove(png_files)
rm(plot_list)

```

# Creat Summary Visual for Best Date in Each State
```{r}
# Estimate state centers
state_centers <- us_states %>%
  group_by(region) %>%
  summarize(lat = min(lat) + (max(lat)-min(lat))/2, 
            long = min(long) + (max(long)-min(long))/2,
            lat2 = NA,
            long2 = NA)

# Adjust for less rectangular-shaped states (me being a perfectionist)
state_centers$lat[state_centers$region=="california"] <- 36.5
state_centers$lat[state_centers$region=="idaho"] <- 43.5
state_centers$lat[state_centers$region=="nevada"] <- 39.5
state_centers$long[state_centers$region=="minnesota"] <- -94.5
state_centers[state_centers$region=="oklahoma", c("lat", "long")] <- data.frame(lat=35.5, long=-97.2)
state_centers$long[state_centers$region=="texas"] <- -99
state_centers$long[state_centers$region=="louisiana"] <- -92.5
state_centers$long[state_centers$region=="florida"] <- -81.5
state_centers[state_centers$region=="south carolina", c("lat", "long")] <- data.frame(lat=33.9, long=-80.5)
state_centers[state_centers$region=="north carolina", c("lat", "long")] <- data.frame(lat=35.5, long=-79)
state_centers[state_centers$region=="virginia", c("lat", "long")] <- data.frame(lat=37.5, long=-78.5)
state_centers[state_centers$region=="west virginia", c("lat", "long")] <- data.frame(lat=38.5, long=-81)
state_centers[state_centers$region=="district of columbia", c("lat2", "long2")] <- state_centers[state_centers$region=="district of columbia", c("lat", "long")]
state_centers[state_centers$region=="district of columbia", c("lat", "long")] <- data.frame(lat=33, long=-68)
state_centers[state_centers$region=="maryland", c("lat", "long")] <- data.frame(lat2=34.5, long2=-68)
state_centers[state_centers$region=="maryland", c("lat2", "long2")] <- data.frame(lat2=39.3, long2=-77)
state_centers[state_centers$region=="delaware", c("lat", "long")] <- data.frame(lat=36, long=-68)
state_centers[state_centers$region=="delaware", c("lat2", "long2")] <- data.frame(lat2=38.8, long2=-75.4)
state_centers[state_centers$region=="new jersey", c("lat2", "long2")] <- data.frame(lat2=40.1, long2=-74.4)
state_centers[state_centers$region=="new jersey", c("lat", "long")] <- data.frame(lat=37.5, long=-68)
state_centers[state_centers$region=="connecticut", c("lat2", "long2")] <- data.frame(lat2=41.6, long2=-72.7)
state_centers[state_centers$region=="connecticut", c("lat", "long")] <- data.frame(lat=39, long=-68)
state_centers[state_centers$region=="rhode island", c("lat2", "long2")] <- data.frame(lat2=41.7, long2=-71.5)
state_centers[state_centers$region=="rhode island", c("lat", "long")] <- data.frame(lat=40.5, long=-68)
state_centers[state_centers$region=="massachusetts", c("lat2", "long2")] <- data.frame(lat2=42.4, long2=-71.7)
state_centers[state_centers$region=="massachusetts", c("lat", "long")] <- data.frame(lat=42.5, long=-68)
state_centers[state_centers$region=="illinois", c("lat", "long")] <- data.frame(lat=40, long=-89.1)
state_centers[state_centers$region=="indiana", c("lat", "long")] <- data.frame(lat=40, long=-86.1)
state_centers[state_centers$region=="michigan", c("lat", "long")] <- data.frame(lat=43, long=-84.5)
state_centers[state_centers$region=="kentucky", c("lat", "long")] <- data.frame(lat=37.4, long=-85.5)
state_centers[state_centers$region=="vermont", c("lat2", "long2")] <- data.frame(lat2=43.9, long2=-72.7)
state_centers[state_centers$region=="vermont", c("lat", "long")] <- data.frame(lat=47, long=-75)
state_centers[state_centers$region=="new hampshire", c("lat2", "long2")] <- data.frame(lat2=43.9, long2=-71.5)
state_centers[state_centers$region=="new hampshire", c("lat", "long")] <- data.frame(lat=47, long=-71.5)

# Add state centers back onto Best Day data
df_BestDay <- df_BestDay %>%
  select(days_into_year, State, Date, Date_short, Region) %>%
  mutate(State = tolower(State)) %>%
  left_join(state_centers, by=c("State"="region"))

# Create the visualization
ggplot() + 
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group), fill="lightgrey", color="black") +
  #geom_point(data = state_centers, aes(x=long, y=lat), color="red", size=1) +
  geom_text(data = df_BestDay, aes(x=long, y=lat, label=Date_short), color="darkblue", size=4, vjust=0.5) + 
  geom_segment(data = state_centers, aes(x=ifelse(long>-70,long-1.4,long), y=ifelse(lat==47,46.4,lat), xend=long2, yend=lat2), color="darkblue") + 
  theme_classic() + 
  labs(title="Best Date to Find Chicken of the Woods Mushroom") + 
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank(),
        plot.title = element_text(hjust=0.5)) + 
  coord_fixed(1.3) # set aspect ratio
```

#References
1. Shackleton, C. M., & de Vos, A. (2022). How many people globally actually use non-timber forest products?. Forest Policy and Economics, 135, 102659. https://doi.org/10.1016/j.forpol.2021.102659

2. Robbins, P., Emery, M., & Rice, J. L. (2008). Gathering in Thoreau's backyard: nontimber forest product harvesting as practice. Area, 40(2), 265-277. https://doi.org/10.1111/j.1475-4762.2008.00794.x